---
title: "Hydrogen-bonded Ionic Liquids (SRS-MP2)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "figs/")
options(width = 140)
require(data.table)
require(magrittr)
require(tidyr)
require(ggplot2)

source("~/GoogleDrive/srs-app/scripts/gen_func.r")
```

The raw geometries of 42 hydrogen-bonded ionic liquids were constructed and then geometry optimised using Truhlar's M06-2X functional along with aug-cc-pVDZ in ethanol. Then CCSD(T)/aVDZ and MP2/(aug)-cc-pVXZ calculations, where X = D,T,Q, were run.

```{r, include = FALSE}
hbil_ccsdt <- read.csv(file = "~/GoogleDrive/srs-app/data/hbil-ccsdt-adz.edat", header = TRUE
         , sep = "|", strip.white = TRUE) %>% data.table()

hbil_mp2 <- read.csv(file = "~/GoogleDrive/srs-app/data/hbil-mp2.edat", header = TRUE
                    , sep = "|", strip.white = TRUE) %>% data.table()

hbil_ccsdt <- gather(hbil_ccsdt, En, value, CorrEn, Total) %>% data.table()            # make super long
hbil_mp2 <- gather(hbil_mp2, En, value, OS:RefEn) %>% data.table()

hbil_ccsdt[, value := 2625.5 * value]                                                 # convert to kJ/mol
hbil_mp2[, value := 2625.5 * value]  

hbil_ccsdt <- spread(hbil_ccsdt, Component, value) %>% data.table()                    # widen by Component
hbil_mp2 <- spread(hbil_mp2, Component, value) %>% data.table()

# A-alone etc. not valid column names
setnames(hbil_ccsdt, names(hbil_ccsdt), make.names(names(hbil_ccsdt)))
setnames(hbil_mp2, names(hbil_mp2), make.names(names(hbil_mp2)))

hbil_ccsdt[, nonCP := complex - A.alone - B.alone]                   # non-counterpoise corrected
hbil_ccsdt[, CP := complex - A.ghostB - B.ghostA]                    # counterpoise corrected

hbil_mp2[, nonCP := complex - A.alone - B.alone]
hbil_mp2[, CP := complex - A.ghostB - B.ghostA]

hbil_ccsdt <- gather(hbil_ccsdt, Component, value, A.alone:CP) %>%
    spread(En, value) %>%
    data.table()

hbil_mp2 <- gather(hbil_mp2, Component, value, A.alone:CP) %>%
    spread(En, value) %>%
    data.table()


setkey(hbil_ccsdt, Suite, Method, Basis, System, Component)
setkey(hbil_mp2, Suite, Method, Basis, System, Component)

# merge CCSD(T) [CorrEn, Total] and MP2 [OS, SS, RefEn] data for aVDZ
# this is for calculating the MP2 correction term for CCSD(T)
# Note: all counterpoise corrected correlation energies (OS + SS)
# should be 10 rows only!
hbil.adz_both.cp <- merge(hbil_ccsdt[Component == "CP", !c("Basis", "Method"), with = F]
                  , hbil_mp2[Basis == "adz" & Component == "CP", !c("Basis", "Method"), with = F]
                  , by = c("Suite", "System", "Component"))

hbil.adz_both.cp[, d_mp2 := CorrEn - OS - SS]

print(hbil.adz_both.cp)

hbil.mp2_cbs <- gather(hbil_mp2, En, value, OS, SS, RefEn) %>%
    spread(Basis, value) %>%
    data.table()

# ================ Helgaker extrapolation =============== #
hbil.mp2_cbs[, cbs := (3^3 * atz - 4^3 * aqz) / (3^3 - 4^3)]

hbil.mp2_cbs <- gather(hbil.mp2_cbs[Component == "CP"], Basis, value, adz:cbs) %>%
    spread(En, value) %>%
    data.table()

print(hbil.mp2_cbs)

hbil.ccsdt_cbs <- merge(hbil.mp2_cbs[(Basis == "cbs"), !c("Method", "Component"), with = F]
                   , hbil.adz_both.cp[, !c("OS", "SS", "RefEn", "Component"), with = F], by = c("Suite", "System"))
hbil.ccsdt_cbs[, ccsdt.cbs := OS + SS + d_mp2]

print(hbil.ccsdt_cbs[order(Basis, System)])
```

And the number of high and low ratios by basis set, as well as the distributions (note the low number $\epsilon_{\Delta s} < 1$ systems):

```{r}
hbil.mp2_spin <- hbil_mp2[Component == "nonCP", !c("Method", "RefEn"), with = FALSE]

hbil.mp2_spin <- merge(hbil.mp2_spin, hbil.ccsdt_cbs[Basis == "cbs", !c("OS", "SS", "RefEn", "Basis", "Total"), with = F]
                  , by = c("Suite", "System"))

hbil.mp2_spin[, eps := OS / SS]
hbil.mp2_spin[, r := as.factor(ifelse(eps > 1, "A", "B"))]

print(hbil.mp2_spin[, nrow(.SD), by = .(r, Basis)])

ggplot(hbil.mp2_spin, aes(x = eps, y = ..count..)) +
    #geom_freqpoly(binwidth = 0.005, aes(colour = Basis, linetype = r)) #+
    geom_density(aes(fill = r), alpha = 0.4) +
    theme_bw() +
    facet_wrap(~ Basis) +
    scale_fill_manual(name = bquote(epsilon[Delta*s]), values = c("orange", "blue")) 

```

The procedure followed is pretty much identical to that of the stacked base pairs, so very briefly:

1. Helgaker extrapolation, using correction term assuming basis set indifference between MP2 and CCSD(T) to obtain CCSD(T)/CBS.
2. All energies used for the benchmark energies were counterpoise corrected.
3. **Non-counterpoise corrected** energies used for SRS-MP2, the coefficients are displayed in the code below.

```{r}
paramA <- data.table(r = as.factor("A")
                    , Basis = c("dz", "tz", "qz", "adz", "atz", "aqz")
                    , c.OS = 0
                    , c.SS = c(1.752, 1.640, 1.689, 1.372, 1.443, 1.591))

paramB <- data.table(r = as.factor("B")
                    , Basis = c("dz", "tz", "qz", "adz", "atz", "aqz")
                    , c.OS = c(0.543, 0.660, 0.671, 0.856, 0.635, 0.699)
                    , c.SS = c(1.518, 1.140, 1.119, 1.033, 1.047, 1.075))

params_SRS.1 <- rbind(paramA, paramB)

# merge 
hbil.srs.dt <- merge(hbil.mp2_spin, params_SRS.1, by = c("Basis", "r"))


hbil.srs.dt[, srs.1 := c.OS * OS + c.SS * SS]
```


And the errors, by basis set:
```{r}
print(hbil.srs.dt[, as.list(nomed.stats(ccsdt.cbs - srs.1)), by = .(Basis)])
```



