---
title: "Stacked Base Pairs"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width = 140)
require(data.table)
require(magrittr)
require(tidyr)

source("~/GoogleDrive/srs-app/scripts/gen_func.r")
```

Here, we have re-optimised the geometries for the stacked base pairs using Truhlar's M06-2X functional along with aug-cc-pVDZ in ethanol. Then CCSD(T)/aVDZ and MP2/(aug)-cc-pVXZ calculations, where X = D,T,Q, were run.

```{r, include = FALSE}
sbp_ccsdt <- read.csv(file = "~/GoogleDrive/srs-app/data/sbp-ccsdt-adz.edat", header = TRUE
         , sep = "|", strip.white = TRUE) %>% data.table()

sbp_mp2 <- read.csv(file = "~/GoogleDrive/srs-app/data/sbp-mp2.edat", header = TRUE
                    , sep = "|", strip.white = TRUE) %>% data.table()

sbp_ccsdt <- gather(sbp_ccsdt, En, value, CorrEn, Total) %>% data.table()            # make super long
sbp_mp2 <- gather(sbp_mp2, En, value, OS:RefEn) %>% data.table()

sbp_ccsdt[, value := 2625.5 * value]                                                 # convert to kJ/mol
sbp_mp2[, value := 2625.5 * value]  

sbp_ccsdt <- spread(sbp_ccsdt, Component, value) %>% data.table()                    # widen by Component
sbp_mp2 <- spread(sbp_mp2, Component, value) %>% data.table()

# A-alone etc. not valid column names
setnames(sbp_ccsdt, names(sbp_ccsdt), make.names(names(sbp_ccsdt)))
setnames(sbp_mp2, names(sbp_mp2), make.names(names(sbp_mp2)))
```


After reading in our data, converting it to kJ/mol, etc. we calculate the counterpoise-corrected and non-counterpoise corrected values for both CCSD(T) and MP2, for all components. In particular, that means correlation energy (CorrEn) for CCSD(T), and OS and SS for MP2. 

```{r, echo=TRUE}
sbp_ccsdt[, nonCP := complex - A.alone - B.alone]                   # non-counterpoise corrected
sbp_ccsdt[, CP := complex - A.ghostB - B.ghostA]                    # counterpoise corrected

sbp_mp2[, nonCP := complex - A.alone - B.alone]
sbp_mp2[, CP := complex - A.ghostB - B.ghostA]

# print(sbp_ccsdt)
# print(sbp_mp2)
```


```{r, include=FALSE}
# back to original shape
sbp_ccsdt <- gather(sbp_ccsdt, Component, value, A.alone:CP) %>%
    spread(En, value) %>%
    data.table()

sbp_mp2 <- gather(sbp_mp2, Component, value, A.alone:CP) %>%
    spread(En, value) %>%
    data.table()
```

Then we merge CCSD(T)/aVDZ and MP2/aVDZ into `adz_both.cp` in order to determine the correction term. Note that counterpoise corrected energies were used!
```{r}
setkey(sbp_ccsdt, Suite, Method, Basis, System, Component)
setkey(sbp_mp2, Suite, Method, Basis, System, Component)

# merge CCSD(T) [CorrEn, Total] and MP2 [OS, SS, RefEn] data for aVDZ
# this is for calculating the MP2 correction term for CCSD(T)
# Note: all counterpoise corrected correlation energies (OS + SS)
# should be 10 rows only!
adz_both.cp <- merge(sbp_ccsdt[Component == "CP", !c("Basis", "Method"), with = F]
                  , sbp_mp2[Basis == "adz" & Component == "CP", !c("Basis", "Method"), with = F]
                  , by = c("Suite", "System", "Component"))
# print(adz_both.cp)
```

After that, we calculate the MP2 correction, the difference between CCSD(T) and MP2 for the aug-cc-pVDZ basis set.

```{r}
adz_both.cp[, d_mp2 := CorrEn - OS - SS]

print(adz_both.cp)
```
Hmm...the `d_mp2` values are all positive!

```{r, include = FALSE}

# to calculate MP2/CBS
mp2_cbs <- gather(sbp_mp2, En, value, OS, SS, RefEn) %>%
    spread(Basis, value) %>%
    data.table()
```

Then, we use the Helgaker extrapolation to calculate MP2/CBS:
```{r}
# ================ Helgaker extrapolation =============== #
mp2_cbs[, cbs := (3^3 * atz - 4^3 * aqz) / (3^3 - 4^3)]
```

Following this, we are only interested in the counterpoise corrected MP2/CBS values for the OS and SS. So select only counterpoise corrected values, and collect the basis sets in a column, while spreading the original `OS`, `SS`, etc. out into columns.

```{r}
mp2_cbs <- gather(mp2_cbs[Component == "CP"], Basis, value, adz:cbs) %>%
    spread(En, value) %>%
    data.table()

print(mp2_cbs)
```

Thus, we merge the MP2/CBS counterpoise corrected values with the MP2 correction term from `adz_both.cp`. Note that the **OS** and **SS** come from counterpoise corrected MP2 (for both aVDZ and CBS), and **`d_mp2`** comes from `CCSD(T)/aVDZ - MP2/aVDZ`, counterpoise corrected. Are my formulas correct?

```{r}
# note using "CBS" basis set, removing "old" and duplicate columns from adz_both.cp
ccsdt_cbs <- merge(mp2_cbs[(Basis == "cbs" | Basis == "adz"), !c("Method", "Component"), with = F]
                   , adz_both.cp[, !c("OS", "SS", "RefEn", "Component"), with = F], by = c("Suite", "System"))
ccsdt_cbs[, ccsdt.cbs := OS + SS + d_mp2]

print(ccsdt_cbs[order(Basis, System)])
```

Now, notice that I used both MP2/CBS and MP2/aVDZ. Here's the weird thing. Below are the old CCSD(T)/CBS results, along with the SRS/VTZ and VQZ that Dennis calculated. The aVDZ results above are much closer to the SRS results below! Whereas the new CBS results are rather high.

```
	ccsd(t)/cbs 	Eint(srs)/tz	Eint(srs)/qz
AA	-52.01886741	-57.09531123	-57.13304834
AC	-46.05043149	-48.30771951	-48.71215329
AU	-42.66938187	-48.40978639	-48.78524818
CC	-38.13748959	-37.28317451	-35.6200886
CG	-40.28430518	-49.9374354	    -44.8954283
CU	-37.50967724	-32.36791902	-33.08033109
GA	-52.30441493	-56.62308196	-56.81254076
GG	-53.99221117	-56.43450303	-56.5931341
GU	-41.20755022	-45.9184363	    -43.39385213
UU	-34.76382056	-34.11152255	-32.34332397
```

Calculating $ \epsilon_{\Delta s}$: 
```{r}
mp2_spin <- sbp_mp2[Component == "nonCP", !c("Method", "RefEn"), with = FALSE]

mp2_spin <- merge(mp2_spin, ccsdt_cbs[Basis == "cbs", !c("OS", "SS", "RefEn", "Basis", "Total"), with = F]
                  , by = c("Suite", "System"))

mp2_spin[, eps := OS / SS]
mp2_spin[, r := as.factor(ifelse(eps > 1, "A", "B"))]
```

Inputting the scaling parameters, and calculating the scaled energies:
```{r}
paramA <- data.table(r = as.factor("A")
                    , Basis = c("dz", "tz", "qz", "adz", "atz", "aqz")
                    , c.OS = 0
                    , c.SS = c(1.752, 1.640, 1.689, 1.372, 1.443, 1.591))

paramB <- data.table(r = as.factor("B")
                    , Basis = c("dz", "tz", "qz", "adz", "atz", "aqz")
                    , c.OS = c(0.543, 0.660, 0.671, 0.856, 0.635, 0.699)
                    , c.SS = c(1.518, 1.140, 1.119, 1.033, 1.047, 1.075))

params_SRS.1 <- rbind(paramA, paramB)

# merge 
srs.dt <- merge(mp2_spin, params_SRS.1, by = c("Basis", "r"))


srs.dt[, srs.1 := c.OS * OS + c.SS * SS]
```

Here are the energies of SRS/aVTZ, compared with CCSD(T)/CBS (new geometries):
```{r}
srs.dt[Basis == "tz", c("System", "ccsdt.cbs", "srs.1"), with = F] %>% print()
```

Now for the moment of truth, the errors:
```{r}
print(srs.dt[, as.list(nomed.stats(ccsdt.cbs - srs.1)), by = .(Basis)])
```

Here are the errors for the three best performing basis sets:
```{r}
srs.dt[Basis == "tz" | Basis == "qz" | Basis == "aqz"
                    , ccsdt.cbs - srs.1, by = .(System, Basis)] %>%
    spread(Basis, V1) %>%
    print()
```

